### 2024-07-15 evaluating-prediction-accuracy.qmd
20240714 - edits

- http://araw.mede.uic.edu/cgi-bin/testcalc.pl
- https://dlrs.shinyapps.io/shinyDLRs/

### Confusion Matrix aka 2x2 Accuracy Table aka Cross-Tabulation aka Contingency Table {#sec-confusionMatrix}

@@

#### Number {#sec-confusionMatrix-number}

@@

#### Number with margins added {#sec-confusionMatrix-numberMargins}

@@

#### Proportions {#sec-confusionMatrix-numberMarginsProportions}

@@

#### Proportions with margins added {#sec-confusionMatrix-numberMarginsProportionsMargins}

@@

### True Positives (TP) {#sec-truePositive}

@@

### True Negatives (TN) {#sec-trueNegative}

@@

### False Positives (FP) {#sec-falsePositive}

@@

### False Negatives (FN) {#sec-falseNegative}

@@

$$
\begin{aligned}
  \text{FN} &= \text{BR} \times (1 - \text{SR}) \times N
\end{aligned}
(\#eq:falseNegative)
$$ {#eq-falseNegative}

### Selection Ratio (SR) {#sec-selectionRatio}

@@

$$
\begin{aligned}
  \text{SR} &= P(R_i) \\
  &= \frac{\text{TP} + \text{FP}}{N}
\end{aligned}
(\#eq:selectionRatio)
$$ {#eq-selectionRatio}

### Base Rate (BR) {#sec-pretestProbability}

@@

### Pretest Odds {#sec-pretestOdds}

@@

$$
\begin{aligned}
  \text{pretest odds} &= \frac{\text{pretest probability}}{1 - \text{pretest probability}} \\
\end{aligned}
(\#eq:pretestOdds)
$$ {#eq-pretestOdds}

### Percent Accuracy {#sec-percentAccuracy}

@@

### Percent Accuracy by Chance {#sec-percentAccuracyByChance}

@@

### Percent Accuracy Predicting from the Base Rate {#sec-percentAccuracyPredictingFromBaseRate}

@@

### Relative Improvement Over Chance (RIOC) {#sec-relativeImprovementOverChance}

@@

### Relative Improvement Over Predicting from the Base Rate {#sec-relativeImprovementOverPredictingFromBaseRate}

@@

### Sensitivity (SN) {#sec-sensitivity}

@@

### Specificity (SP) {#sec-specificity}

@@

### False Negative Rate (FNR) {#sec-falseNegativeRate}

@@

### False Positive Rate (FPR) {#sec-falsePositiveRate}

@@

$$
\begin{aligned}
  \text{false positive rate (FPR)} &= P(R|\text{not } C) \\
  &= \frac{\text{FP}}{\text{FP} + \text{TN}} = \frac{\text{FP}}{N (1 - \text{BR})} = 1 - \text{TNR}
\end{aligned}
(\#eq:falsePositiveRate)
$$ {#eq-falsePositiveRate}

### Positive Predictive Value (PPV) {#sec-ppv}

@@

### Negative Predictive Value (NPV) {#sec-npv}

@@

### False Discovery Rate (FDR) {#sec-falseDiscoveryRate}

@@

### False Omission Rate (FOR) {#sec-falseOmissionRate}

@@

### Youden's J Statistic {#sec-youdenJ-example}

@@

### Balanced Accuracy {#sec-balancedAccuracy}

Balanced accuracy is the average of [sensitivity](#sec-sensitivity) and [specificity](#sec-specificity).
Higher values reflect greater accuracy.
The formula for calculating balanced accuracy is in @balancedAccuracy.

@@

### F-Score {#sec-fScore}

@@

### Matthews Correlation Coefficient (MCC) {#sec-matthewsCorrelationCoefficient}

@@

### Diagnostic Odds Ratio {#sec-diagnosticOddsRatio}

@@

### Diagnostic Likelihood Ratio {#sec-diagnosticLikelihoodRatio2}

@@

#### Positive Likelihood Ratio (LR+) {#sec-positiveLikelihoodRatio2}

@@

#### Negative Likelihood Ratio (LR−) {#sec-negativeLikelihoodRatio2}

@@

### Posttest Odds {#sec-posttestOdds}

@@

### Posttest Probability {#sec-posttestProbability}

>>>>

- <http://araw.mede.uic.edu/cgi-bin/testcalc.pl>
- <https://dlrs.shinyapps.io/shinyDLRs>

#### Confusion Matrix aka 2x2 Accuracy Table aka Cross-Tabulation aka Contingency Table {#sec-confusionMatrix}

@@

##### Number {#sec-confusionMatrix-number}

@@

##### Number with margins added {#sec-confusionMatrix-numberMargins}

@@

##### Proportions {#sec-confusionMatrix-numberMarginsProportions}

@@

##### Proportions with margins added {#sec-confusionMatrix-numberMarginsProportionsMargins}

@@

#### True Positives (TP) {#sec-truePositive}

@@

#### True Negatives (TN) {#sec-trueNegative}

@@

#### False Positives (FP) {#sec-falsePositive}

@@

#### False Negatives (FN) {#sec-falseNegative}

@@

$$
\begin{aligned}
  \text{FN} &= \text{BR} \times (1 - \text{SR}) \times N
\end{aligned}
$$ {#eq-falseNegative}

#### Selection Ratio (SR) {#sec-selectionRatio}

@@

$$
\begin{aligned}
  \text{SR} &= P(R_i) \\
  &= \frac{\text{TP} + \text{FP}}{N}
\end{aligned}
$$ {#eq-selectionRatio}

#### Base Rate (BR) {#sec-pretestProbability}

@@

#### Pretest Odds {#sec-pretestOdds}

@@

$$
\begin{aligned}
  \text{pretest odds} &= \frac{\text{pretest probability}}{1 - \text{pretest probability}} \\
\end{aligned}
$$ {#eq-pretestOdds}

#### Percent Accuracy {#sec-percentAccuracy}

@@

#### Percent Accuracy by Chance {#sec-percentAccuracyByChance}

@@

#### Percent Accuracy Predicting from the Base Rate {#sec-percentAccuracyPredictingFromBaseRate}

@@

#### Relative Improvement Over Chance (RIOC) {#sec-relativeImprovementOverChance}

@@

#### Relative Improvement Over Predicting from the Base Rate {#sec-relativeImprovementOverPredictingFromBaseRate}

@@

#### Sensitivity (SN) {#sec-sensitivity}

@@

#### Specificity (SP) {#sec-specificity}

@@

#### False Negative Rate (FNR) {#sec-falseNegativeRate}

@@

#### False Positive Rate (FPR) {#sec-falsePositiveRate}

@@

$$
\begin{aligned}
  \text{false positive rate (FPR)} &= P(R|\text{not } C) \\
  &= \frac{\text{FP}}{\text{FP} + \text{TN}} = \frac{\text{FP}}{N (1 - \text{BR})} = 1 - \text{TNR}
\end{aligned}
$$ {#eq-falsePositiveRate}

#### Positive Predictive Value (PPV) {#sec-ppv}

@@

#### Negative Predictive Value (NPV) {#sec-npv}

@@

#### False Discovery Rate (FDR) {#sec-falseDiscoveryRate}

@@

#### False Omission Rate (FOR) {#sec-falseOmissionRate}

@@

#### Youden's J Statistic {#sec-youdenJ-example}

@@

#### Balanced Accuracy {#sec-balancedAccuracy}

Balanced accuracy is the average of [sensitivity](#sec-sensitivity) and [specificity](#sec-specificity).
Higher values reflect greater accuracy.
The formula for calculating balanced accuracy is in @eq-balancedAccuracy.

@@

#### F-Score {#sec-fScore}

@@

#### Matthews Correlation Coefficient (MCC) {#sec-matthewsCorrelationCoefficient}

@@

#### Diagnostic Odds Ratio {#sec-diagnosticOddsRatio}

@@

#### Diagnostic Likelihood Ratio {#sec-diagnosticLikelihoodRatio2}

@@

##### Positive Likelihood Ratio (LR+) {#sec-positiveLikelihoodRatio2}

@@

##### Negative Likelihood Ratio (LR−) {#sec-negativeLikelihoodRatio2}

@@

#### Posttest Odds {#sec-posttestOdds}

@@

#### Posttest Probability {#sec-posttestProbability}
